A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-07-31 paper/elsarticle/elsarticle-template.tex
paper

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen. \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display. Care must be taken in carrying out spatial AR, however, to warp the images to be projected in a manner that they appear free of distortions to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene. For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.

@@

Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen. Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display. Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to partake in the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive. In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems. Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions. Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.

@@

An early spatial AR system explicitly using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinharez \cite{pinhanez2001everywhere}. For the purposes of the prototype, a projector mounted with a steerable mirror was set up with a camera to demonstrate a variety of use cases, including collaborative assembly encompassing the projection of assembly instructions and an interactive projected user interface \cite{kjeldsen2002interacting,pinhanez2003applications}. The authors, however, carry out keystone correction manually, by interactively adjusting for the position and orientation of a virtual scene plane and a scaling factor and computing a 2D homography accordingly. %Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}. Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to rigid motion.

More generally, keystone correction for planar scenes reduces to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours. The intuition for why it is that a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity'). To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines would be sufficient.

One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface. The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane. Tilt sensors can be used to recover the projector's gravity vector, and be integrated in the computation of a homography as in Raskar and Beardsley \cite{raskar2001self}.

@@

@@

Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$ that minimize the cumulative reprojection~error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ of the point~$\mathbf{X}_{i,j}$ to that image. The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}. The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.

Calibration of a projector can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences in support of calibrating cameras and projectors. A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the scene plane.

@@

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively relative to the placement of the camera.

>>>>

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen. \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display. Care must be taken in carrying out spatial AR, however, to warp the images to be projected in a manner that they appear free of distortions to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene. For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process referred to as keystone correction, often using software bundled with the projector.

@@

Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards,uva2018evaluating,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen. Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display. Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to partake in the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive. In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems. Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions. Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.

@@

An early spatial AR system explicitly using a projector mounted with a steerable mirror is the IBM Everywhere Displays prototype of Pinharez \cite{pinhanez2001everywhere}. For the purposes of the prototype, a projector mounted with a steerable mirror was set up with a camera to demonstrate a variety of use cases, including collaborative assembly encompassing the projection of assembly instructions and an interactive projected user interface \cite{kjeldsen2002interacting,pinhanez2003applications}. The authors, however, carry out keystone correction manually, by interactively adjusting for the position and orientation of a virtual scene plane and a scaling factor and computing a 2D homography accordingly. %Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}. Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to rigid motion.

More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours. The intuition for why it is that a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity'). To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines is sufficient.

One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface \cite{Hartley2004}. The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane. Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector, which they use to carry out a rotational correction in projecting to a wall.

@@

Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$ that minimize the cumulative reprojection~error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ of the point~$\mathbf{X}_{i,j}$ to that image. The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}. The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{}) \in SE(3)$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern.

Calibration of a projector can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene. What remains of this section is concerned primarily with the recovery of 2D-3D correspondences in support of calibrating cameras and projectors. A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.

@@

@@

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively in accordance with the placement of the camera.
